Cross-lingual Document Similarity and Event Detection \todo{Better title...}

### 2015-02-21 rupnik2015.tex
- added learning features

One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as "any significant happening that is being reported about in the media". Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

@@

@@

>>>>

One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

@@

In order to use a supervised approach to determine if two clusters $c_i$ and $c_j$ are equivalent we need to extract from the cluster pair as many informative features as possible. Based on the availability of the data, we have extracted the following groups of features:

\begin{itemize}
\item {\bf CCA related features}. CCA was already used to obtain candidates for the equivalent clusters. Additionally, it can also be used to provide valuable learning features. There are two main features that we extract using CCA -- \texttt{linkCount} and \texttt{avgSimScore}. \texttt{linkCount} is the number of times an article $a_i \in c_i$ has as one of the 10 most similar articles an article $a_j \in c_j$. In other words, it is the number of times an article from $c_i$ has a very similar article, which is in $c_j$. Beside the number of links between the two clusters we can also take into account the similarity scores of the links. The \texttt{avgSimScore} is the feature that represents the average similarity score of the links between the two clusters.

\item {\bf Concept related features}. As mentioned before, articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page (\emph{http://en.wikipedia.org/wiki/Barack\_Obama}). In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, we can almost in all cases normalize the Wikipedia links to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although articles are in different languages, the annotations can therefore provide a shared vocabulary that can be used to compare articles/clusters. By analyzing all articles in clusters $c_i$ and $c_j$ we can identify the most relevant entities and keywords for each cluster. Additionally we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights we can now define a number of features. \texttt{entityCosSim} represents the cosine similarity on the vector of entities mentioned in both clusters. Similarly, \texttt{keywordCosSim} represents the cosine similarity on the keywords only. In case the weights are not relevant, we have also added two features (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By analyzing the locations mentioned in the articles in a cluster we are able to estimate where the event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For articles in each cluster we first compute the average date and time when the articles were published. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The last feature in this group is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the ratio of all dates that are detected in both clusters.

\end{itemize}

@@

In order to evaluate the quality of individual features and the overall model we have performed a set of experiments.
